---
title: "Project 1"
output: html_document
date: "2023-09-12"
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Markdown

## **Packages**

```{r}
library(tidyverse)
library(dplyr)
library(tidyr)
library(forcats)
library(ggplot2)
library(gapminder)
library(stringr)

```

## Downloading Data

```{r}
tuesdata <- tidytuesdayR::tt_load("2022-01-18")
chocolate <- tuesdata$chocolate
as_tibble(chocolate)
glimpse(chocolate)
```

## Part 1: Exploring Data

1.  Make a histogram of the `rating` scores to visualize the overall
    distribution of scores. Change the number of bins from the default
    to 10, 15, 20, and 25. Pick on the one that you think looks the
    best. Explain what the difference is when you change the number of
    bins and explain why you picked the one you did.

```{r}
qplot(rating, data=chocolate, bins = 10)
qplot(rating, data=chocolate, bins = 15)
qplot(rating, data=chocolate, bins = 20)
qplot(rating, data=chocolate, bins = 25)

#The one that looks the best
qplot(rating, data=chocolate, bins = 25)
```

Changing the bin number resulted in a separation the data into more
categories. In essence, a higher bin number allowed for more categories
of ratings, and subsequently, spread out the ratings in the histogram. I
chose the graph with a bin number of 25 because it allowed me to see
more of the data, since the bins themselves were smaller and there were
more of them. I also liked the aesthetic look of the plot the best.

2.  Consider the countries where the beans originated from. How many
    reviews come from each country of bean origin?

```{r}
chocolate %>%
  group_by(country_of_bean_origin) %>%
  count(chocolate$country_of_bean_origin)
```

3.  What is average `rating` scores from reviews of chocolate bars that
    have Ecuador as `country_of_bean_origin` in this dataset? For this
    same set of reviews, also calculate (1) the total number of reviews
    and (2) the standard deviation of the `rating` scores. Your answer
    should be a new data frame with these three summary statistics in
    three columns. Label the name of these columns `mean`, `sd`, and
    `total`.

```{r}

chocolate.f <- filter(chocolate, country_of_bean_origin == "Ecuador")
chocolate.subset <- select(chocolate.f, rating)

chocolate.subset <- mutate(chocolate.subset, mean = mean(rating, na.rm=TRUE))
chocolate.subset <- mutate(chocolate.subset, total = count(chocolate.f))
chocolate.subset <- mutate(chocolate.subset, sd = sd(rating, na.rm = TRUE))
head(chocolate.subset)

```

4.  Which company (name) makes the best chocolate (or has the highest
    ratings on average) with beans from Ecuador?

```{r}
company_avg_rating <-chocolate.f %>%
  group_by(company_manufacturer) %>%
  summarize(avg_rating = mean(rating, na.rm =TRUE))
head(company_avg_rating %>%
  arrange(desc(avg_rating)), 3)

```

5.  Calculate the average rating across all country of origins for
    beans. Which top 3 countries (for bean origin) have the highest
    ratings on average?

```{r}
avg_rating <- mean(chocolate$rating, na.rm = TRUE)
print(avg_rating)

country_avg_ratings1 <- chocolate %>%
  group_by(country_of_bean_origin) %>%
  summarize(avg_rating = mean(rating, na.rm = TRUE)) %>%
  arrange(desc(avg_rating))
head(select(country_avg_ratings1, avg_rating, country_of_bean_origin), 3)
```

6.  Calculate the average rating across all country of origins for
    beans. Which top 3 countries (for bean origin) have the highest
    ratings on average?

```{r}
str(chocolate)

chocolate_filtered  <- chocolate %>%
  group_by(country_of_bean_origin) %>%
  filter(n() >= 10)

country_avg_ratings2 <- chocolate_filtered %>%
  group_by(country_of_bean_origin) %>%
  summarize(avg_rating = mean(rating, na.rm = TRUE))

head(country_avg_ratings2 %>%
       arrange(desc(avg_rating)), 3)
```

7.  For this last part, let\'s explore the relationship between percent
    chocolate and ratings.

    Use the functions in `dplyr`, `tidyr`, and `lubridate` to perform
    the following steps to the `chocolate` dataset:

    1.  Identify the countries of bean origin with at least 50 reviews.
        Remove reviews from countries are not in this list.

```{r}
chocolate_f50  <- chocolate %>%
  group_by(country_of_bean_origin) %>%
  filter(n() >= 50)

countries_with_50ratings <- chocolate_f50 %>%
  pull(country_of_bean_origin) %>%
  unique()

```

2.  Using the variable describing the chocolate percentage for each
    review, create a new column that groups chocolate percentages into
    one of four groups: (i) \<60%, (ii) \>=60 to \<70%, (iii) \>=70 to
    \<90%, and (iii) \>=90% (**Hint** check out the `substr()` function
    in base R and the `case_when()` function from `dplyr` -- see example
    below).

```{r}
chocolate_f50 <- chocolate_f50 %>%
  mutate(chocolate_perc = case_when(
    substr(cocoa_percent, 1, 2) < 60 ~ "<60%",
    substr(cocoa_percent, 1, 2) >= 60 & substr(cocoa_percent, 1, 2) < 70 ~ ">=60% to <70%",
    substr(cocoa_percent, 1, 2) >= 70 & substr(cocoa_percent, 1, 2) < 90 ~ ">=70% to <90%",
    substr(cocoa_percent, 1, 2) >= 90 ~ ">=90%"
  ))

head(select(chocolate_f50, country_of_bean_origin, cocoa_percent, chocolate_perc))

```

3.  Using the new column described in #2, re-order the factor levels (if
    needed) to be starting with the smallest percentage group and
    increasing to the largest percentage group (**Hint** check out the
    `fct_relevel()` function from `forcats`).

```{r}
chocolate_f50 <- chocolate_f50 %>%
  mutate(chocolate_perc = fct_relevel(chocolate_perc,
                                      "<60%", ">=60% to <70%", ">=70% to <90%", ">=90%"
  ))
```

4.  For each country, make a set of four side-by-side boxplots plotting
    the groups on the x-axis and the ratings on the y-axis. These plots
    should be faceted by country.

```{r}
g <- ggplot(chocolate_f50, aes(
  x= chocolate_perc,
  y= rating
))

g <- chocolate_f50 %>%
  ggplot(aes(chocolate_perc, rating))
print(g)

g +
  geom_boxplot(aes(color = chocolate_perc)) +
  facet_grid(.~ country_of_bean_origin) +
  theme(axis.text.x=element_blank())

```

On average, which category of chocolate percentage is most highly rated?
Do these countries mostly agree or are there disagreements?

On average, the \>= 70% to \<90% category is rated the highest across
all countries with 50 or more ratings. However, not every country rated
this category the highest. For example, Tanzania disagreed and rated the
\>= 60% to \<70% category higher.

## Part 2: Join two datasets together

1.  Use this dataset it to create a new column called `continent` in our
    `chocolate` dataset that contains the continent name for each review
    where the country of bean origin is.

```{r}
country_map <- gapminder %>%
  select(country, continent) %>%
  distinct()

chocolate_part2 <- chocolate %>%
  left_join(x = chocolate, y = country_map, by =c("country_of_bean_origin" ="country"))

chocolate_part2 <-chocolate_part2 %>%
  mutate(continent = case_when(
    (country_of_bean_origin == "Belize"|country_of_bean_origin == "Trinidad"|country_of_bean_origin == "St. Lucia") & is.na(continent) ~ "Americas",
    TRUE ~ continent
  )
  )

chocolate_part2 <-chocolate_part2 %>%
  mutate(continent = case_when(
    (country_of_bean_origin == "Congo"| country_of_bean_origin == "Sao Tome") & is.na(continent) ~ "Africa",
    TRUE ~ continent
  )
  )


chocolate_part2 <-chocolate_part2 %>%
  mutate(continent = case_when(
    (country_of_bean_origin == "Fiji"|country_of_bean_origin == "Papua New Guinea"| country_of_bean_origin == "Solomon Islands" |country_of_bean_origin == "Vanuatu")  & is.na(continent) ~ "Oceania",
    TRUE ~ continent
  )
  )



chocolate_part2 <-chocolate_part2 %>%
  mutate(continent = case_when(
    (country_of_bean_origin == "Grenada" | country_of_bean_origin == "U.S.A.") & is.na(continent) ~ "Americas",
    TRUE ~ continent
  )
  )

na_continent_rows <- subset(chocolate_part2, is.na(continent))

```

2.  Only keep reviews that have reviews from countries of bean origin
    with at least 10 reviews.

```{r}
chocolate_part2 <- chocolate_part2 %>%
  group_by(country_of_bean_origin) %>%
  filter(n() >= 10)
```

3.  Also, remove the country of bean origin named `"Blend"`.

```{r}
chocolate_part2 <- chocolate_part2 %>%
  filter(country_of_bean_origin != "Blend")
```

4.  Make a set of violin plots with ratings on the y-axis and
    `continent`s on the x-axis.

```{r}
g2 <- ggplot(chocolate_part2, aes(x = continent, y = rating))
g2 + geom_violin()
```

## Part 3: Convert wide data into long data 

1.  Create a new set of columns titled `beans`, `sugar`, `cocoa_butter`,
    `vanilla`, `letchin`, and `salt` that contain a 1 or 0 representing
    whether or not that review for the chocolate bar contained that
    ingredient (1) or not (0).

    ```{r}
    chocolate_part3 <- chocolate_part2 %>%
      separate(ingredients, into = c("ingredient_count", "ingredient_list"), sep = "-") %>%
      mutate(
        beans = ifelse(grepl("B", ingredient_list), 1, 0),
        sugar = ifelse(grepl("S", ingredient_list), 1, 0),
        cocoa_butter = ifelse(grepl("C", ingredient_list), 1, 0),
        vanilla = ifelse(grepl("V", ingredient_list), 1, 0),
        lecithin = ifelse(grepl("L", ingredient_list), 1, 0),
        salt = ifelse(grepl("S", ingredient_list), 1, 0)
      ) %>%
      select(-ingredient_count, -ingredient_list)
    view(chocolate_part3)
    ```

2.  Create a new set of columns titled `char_cocoa`, `char_sweet`,
    `char_nutty`, `char_creamy`, `char_roasty`, `char_earthy` that
    contain a 1 or 0 representing whether or not that the most memorable
    characteristic for the chocolate bar had that word (1) or not (0).
    For example, if the word \"sweet\" appears in the
    `most_memorable_characteristics`, then record a 1, otherwise a 0 for
    that review in the `char_sweet` column (**Hint**: check out
    `str_detect()` from the `stringr` package).

    ```{r}
    char_words <- c("cocoa", "sweet", "nutty", "creamy", "roasty", "earthy")

    chocolate_part3 <- chocolate_part3 %>%
      mutate(
        char_cocoa = ifelse(str_detect(most_memorable_characteristics, "cocoa"), 1, 0),
        char_sweet = ifelse(str_detect(most_memorable_characteristics, "sweet"), 1, 0),
        char_nutty = ifelse(str_detect(most_memorable_characteristics, "nutty"), 1, 0),
        char_creamy = ifelse(str_detect(most_memorable_characteristics, "creamy"), 1, 0),
        char_roasty = ifelse(str_detect(most_memorable_characteristics, "roasty"), 1, 0),
        char_earthy = ifelse(str_detect(most_memorable_characteristics, "earthy"), 1, 0)
      )
    view(chocolate_part3)
    ```

3.  For each year (i.e. `review_date`), calculate the mean value in each
    new column you created across all reviews for that year. (**Hint**:
    If all has gone well thus far, you should have a dataset with 16
    rows and 13 columns).

    ```{r}
    mean_by_year <- chocolate_part3 %>%
      group_by(review_date) %>%
      summarize(
        mean_char_cocoa = mean(char_cocoa),
        mean_char_sweet = mean(char_sweet),
        mean_char_nutty = mean(char_nutty),
        mean_char_creamy = mean(char_creamy),
        mean_char_roasty = mean(char_roasty),
        mean_char_earthy = mean(char_earthy),
        mean_beans = mean(beans),
        mean_sugar = mean(sugar),
        mean_cocoa_butter = mean(cocoa_butter),
        mean_vanilla = mean(vanilla),
        mean_lecithin = mean(lecithin),
        mean_salt = mean(salt)
      )
    view(mean_by_year)
    ```

4.  Convert this wide dataset into a long dataset with a new `feature`
    and `mean_score` column.

```{r}
long_choco_part3 <- mean_by_year %>%
  pivot_longer(
    cols = starts_with("mean_"),
    names_to = "feature",
    values_to = "mean_score"
  )

long_choco_part3 <- long_choco_part3 %>%
  mutate(feature = str_replace(feature, "mean_", ""))
long_choco_part3 <- long_choco_part3 %>%
  mutate(feature = str_replace(feature, "char_", ""))

view(long_choco_part3)

```

## **Part 4: Data visualization** 

1.  An overall title for the plot and a subtitle summarizing key trends
    that you found. Also include a caption in the figure with your name.

2.  Both the observed points for the `mean_score`, but also a smoothed
    non-linear pattern of the trend

3.  All plots should be shown in the one figure

4.  There should be an informative x-axis and y-axis label

```{r}
g3 <- ggplot(long_choco_part3, aes(x= review_date,
                                   y=mean_score))

g3 +
  geom_point(aes(color = feature), size = 3, alpha = 1 / 2) +
  geom_smooth(method="lm") +
  labs(
    title= "Average Score of Chocolate Over Time",
    subtitle = "Salty chocolate is increasing over time and sweet chocolate is decreasing",
    caption = "Allison Zeher") +
  labs(
    x = "Date",
    y= "Average Score"
  )
```

## **Part 5: Make the worst plot you can!**

1.  Make the absolute worst plot that you can. You need to customize it
    in **at least 7 ways** to make it awful.

2.  In your document, write 1 - 2 sentences about each different
    customization you added (using bullets -- i.e. there should be at
    least 7 bullet points each with 1-2 sentences), and how it could be
    useful for you when you want to make an awesome data visualization.

```{r}
g3 +
  geom_polygon(color="black", size = 7, alpha = 1/30) +
  labs(title="Chocolate rating") +
  geom_smooth(color="black", linewidth= 0.2, linetype=1) +
  theme_void() +
  labs(x="x",
       y="y") +
  theme(legend.position="top", legend.key.size = unit(.011,"mm"))
```

-   changed the points to polygons so the reader cannot see individual
    data points
-   made the polygons all black and large so that readers cannot
    distinguish features
-   changed the points to 1/40 so it is so light that readers can no
    longer see it
-   changed the title of the plot to something unhelpful with no
    information
-   made the lowess smoother very small
-   made the legend key size so small that readers cannot see the keys
-   changed the theme so that there are no axis tick marks

## **Part 6: Make my plot a better plot!**

1.  You need to customize it in **at least 7 ways** to make it better.

2.  In your document, write 1 - 2 sentences about each different
    customization you added (using bullets -- i.e. there should be at
    least 7 bullet points each with 1-2 sentences), describing how you
    improved it.

```{r}
g3 +
  geom_point(aes(color=feature), size = 3, alpha = 1/2) +
  labs(title="Average Score of Chocolate Over Time") +
  geom_smooth(aes(color=feature), linewidth= 1, se=FALSE, method="lm") +
  theme_minimal() +
  labs(x="Years",
       y="Average Rating") +
  theme(legend.position="right", legend.key.size = unit(2,"mm"))
```

-   made the points color coded so readers can easily see different
    features
-   made the points smaller so readers can accurately see where they are
    located
-   changed the points to 1/2 so readers can see multiple points that
    are close together
-   changed the title of the plot to explain what is being displayed
-   removed the standard errors so that the graph is less cluttered
-   moved the legend key to a location that is easy to find
-   changed the theme to minimal to make the graph more readable
